---
title: "Travelling Salesperson"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: 3
    theme: readable
    includes:
      after_body: footer.html
      in_header: header.html
---
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px;}
</style>
### Objective
An abstract for our project is available on our [home page](index.html). On this page we are going to discuss
the R simulation that computes the mean travel time of 
$m+1$ circuits identically distributed uniformly at random. $m$ and $n$ will be parameters to our simulation where $n$ is number of trials until equilibrium from which the expectation of the next $m$ can be worked out in this fashion,  $$ \sum_{i} f(i)\mu(i) \approx \frac{1}{m+1} \sum_{k=0}^{m} f(X_k)$$ 
<br>
Towards this goal we first label 20 cities numerically from 1:20
```{r}
U <-c(1:20) #vertices
```
In graph terminology these are our vertices and will make up the elements within our state space. Next we resolve the issue of forbidden travel and the missing component of our Hamiltonian graph, edges.  

### Distance Matrix
```{r}
D <- matrix(runif(n=20*20, min=0, max =1), ncol = 20) #uniformly distributed distance matrix
```
We then manually specify 5 pairs $(i,j)$, $i,j \in U$ for which travel is forbidden(i.e. the edge between $i$, $j$ is weighted 0). In the context of our simulation that means a $0$ entry in the distance matrix. 
```{r}
D[4,6] = 0
D[9,13]= 0
D[1,19]= 0 
D[10,17]= 0  
D[3,8] = 0 #NO TRAVEL NO DISTANCE
```
We are almost done setting up our graph except for a couple outstanding requirements.First the edge connecting a pair of cities is more than likely not identical.Second, travel from city $i$ to city $i$ for $i \in U$ is greater than $0$ and most importantly we still haven't assigned the other half of the travel restrictions. To satisfy these requirement, we transform $D$ into a symmetric matrix by assigning the lower triangular of $D$ the entries of the upper triangular of the transpose of D and explicitly set the diagonal entries to zero. Thus as you can see there are 185 pairs with positive real distance,  $(19*20)/2 - 5$ from observing $D$ or ${20 \choose 2}-5$. 
```{r}
D[lower.tri(D)] = t(D)[lower.tri(D)] #symmetric matrix
diag(D) <- rep(0,20) #diagonals are 0
```


Notice that the forbidden pairs were [entered](#distance-matrix) in the form of D[i,j] where i<j or else when we go to make the matrix symmetric the upper triangular entry is still non-zero. This can be remedied by manually making all 10 entries 0, or specifying D[i,j], j<i and assigning the upper-triangular of D with the upper-triangular of the transpose of D
```{r}
D[upper.tri(D)] = t(D)[upper.tri(D)] #symmetric matrix
```

or in a more proper application handling user input.

<br>

### Algorithm 

Thus with the vertices set up and the weighted-edges specified, we have initialized our graph. We can now consider our $Q$. As mentioned above, the most prudent idea as suggested is a method which picks two points uniformly at random and interchanges them. The code chunck below is portion of the function that accomplishes this task and our broader [objective](#objective).
```{r eval=FALSE}
tspMCMC <- function(n,m){
v <- sample(U,size = 20, replace=FALSE) #without repetition sample first sequence
while(indicatorPermissible(v)==0){#make sure its valid
  v <- sample(U,size=20, replace=FALSE) 
}
S <- list(v) #record it in the state space
...
...
...
while(i <= m+n){
  P <- sample(U,size=1) #basic sampling 
  Q <- sample(U,size=1) #likewise for proposed state, 
  index_p = match(P, unlist(S[i])) #the position of P in the sequence
  index_q = match(Q, unlist(S[i])) #likewise for the second element in the pair
  k <- unlist(S[i])  
  z <- c(replace(k, c(index_p,index_q), c(Q,P))) #interchange P and Q 
  if(indicatorPermissible(z) == 1){#if the new sequence is valid
    if(i >= n){#has equilibrium been reached yet?
      travel_time[j] = distance_travelled(S[i]) #store travel time
      j <- j + 1
    }
    S[i+1] <- list(z) #adds the new sequence to collection of sampled state spaces
    i <- i + 1 
  }
  else{
    if(i<=n){#this is done to make sure we dont double count travel times
      S[i+1] <- S[i]
      i <- i + 1
    }
    rejection_counter = rejection_counter+1 #see how many time our proposed transition introduced forbidden travel
  }
}
}
```
There is alot to unpack but very concisely it takes two arguments $n$(integer) and $m$(integer). We then randomly samples a sequence without-repetition which we check is valid and add to our record of the sampled sequence. From there we sample two points, extracts their place in the sequence and use the last record valid sequence, $X_{i}$ to propose a new sequence. Since we are using a Metropolis with $\mu = 1$(i.e. no rejection) we are only responsible for checking the sequence $X_{i+1}$ is valid. This is where *indicatorPermissable($X_{i+1}$)* comes in. It's the responsibility of the *indicatorPermissable* function to return $0$ if the proposed interchange creates a sequence where travel between two forbidden cities is introduced. In the case where the indicator returns $1$ we accept the new sequence as valid and add it to the record(i.e *list*) of valid circuits and increment $i[1]$.Furthermore if we have already reached equilibrium or $n$ trials, we also start computing the travel times by passing the sequence to a function called *distance_travelled*. 
<div class="blue">
[1]As you can see in the code chunk above, our control sequence of choice is a while loop. The reason being, unlike a for-loop we have more control on our index $i$.For example, until the simulation reaches equilibrium we aren't entirely concerned with repeat sequences. When we go to compute travel-times($i \geq n$) however we control for repeat samples by ignoring trials for which *indicatorPermissable* returns 0 and trying again at index $i$. 
</div>

<br>

### Observations
Now let us test the algorithm using the same arbitrary forbidden cities from above and a $Uni(0,1)$ distributed distance matrix, additionally choose $n=1000$ and $m=5000$ 
```{r eval=FALSE}
tspMCMC(1000,5000)

```

```{r echo=FALSE}
U <-c(1:20) #vertices
D <- matrix(runif(n=20*20, min=0, max =1), ncol = 20) #non-stochastic uniformly distributed distance matrix 
D[4,6] = 0
D[9,13]= 0
D[1,19]= 0 
D[10,17]= 0  
D[3,8] = 0 #NO TRAVEL NO DISTANCE, REMEMBER FORBIDDEN PAIRS MUST FOLLOW D[i,j]=0, i<j
D[lower.tri(D)] = t(D)[lower.tri(D)] #symmetric matrix
diag(D) <- rep(0,20) #diagonals are 0

indicatorPermissible <- function(z){
  if(D[z[1],z[20]] == 0){ #checks the endpoints of the sequence
    return(0)
  }
  for(i in 1:19){ #looking at the vertex at index i, consults D to make sure the edge connecting to the vertex at index i+1 is non-zero
    if(D[z[i],z[i+1]] == 0)
      return(0)
  }
  return(1)
}

distance_travelled <- function(L){
  C <- unlist(L)
  f <- D[C[20],C[1]] #endpoints require special care
  for(i in 1:19){
    f = f + D[C[i],C[i+1]]  #similar to indicatorPermissible, except it tallys the edges
  }
  return(f)
}

tspMCMC <- function(n,m){
v <- sample(U,size = 20, replace=FALSE) #without repetition sample first sequence
while(indicatorPermissible(v)==0){#make sure its valid
  v <- sample(U,size=20, replace=FALSE) 
}
S <- list(v) #record it in the state space
rejection_counter <- 0
trials <- 1
i <- 1
j <- 1
repetition_counter <- 0
travel_time <- rep(0, m+1) #init. the travel time vector
while(i <= m+n){
  P <- sample(U,size=1) #basic sampling 
  Q <- sample(U,size=1) #likewise for proposed state, 
  index_p = match(P, unlist(S[i])) #the position of P in the sequence
  index_q = match(Q, unlist(S[i])) #likewise for the second element in the pair
  k <- unlist(S[i])  
  z <- c(replace(k, c(index_p,index_q), c(Q,P))) #interchange P and Q 
  if(indicatorPermissible(z) == 1){#if the new sequence is valid
    if(i >= n){#has equilibrium been reached yet?
      travel_time[j] = distance_travelled(S[i]) #store travel time
      j <- j + 1
    }
    S[i+1] <- list(z) #adds the new sequence to collection of sampled state spaces
    i <- i + 1 
  }
  else{
    if(i<=n){#this is done to make sure we dont double count travel times
      S[i+1] <- S[i]
      i <- i + 1
    }
    rejection_counter = rejection_counter+1 #see how many time our proposed transition introduced forbidden travel
  }
}
   answer <- mean(travel_time)
   return(answer)
}
tspMCMC(1000,5000)
```
Furthermore re-initializing our graph but keeping the distance matrix fixed we can observe the results for multiple arbitrary choices of n and m.
```{r eval=FALSE}
tspMCMC(1000,5000)
tspMCMC(2000,10000)
tspMCMC(4000,20000)
tspMCMC(60000,300000)
```

```{r echo=FALSE}
U <-c(1:20) #vertices
D <- matrix(runif(n=20*20, min=0, max =1), ncol = 20) #non-stochastic uniformly distributed distance matrix 
D[4,6] = 0
D[9,13]= 0
D[1,19]= 0 
D[10,17]= 0  
D[3,8] = 0 #NO TRAVEL NO DISTANCE, REMEMBER FORBIDDEN PAIRS MUST FOLLOW D[i,j]=0, i<j
D[lower.tri(D)] = t(D)[lower.tri(D)] #symmetric matrix
diag(D) <- rep(0,20) #diagonals are 0

indicatorPermissible <- function(z){
  if(D[z[1],z[20]] == 0){ #checks the endpoints of the sequence
    return(0)
  }
  for(i in 1:19){ #looking at the vertex at index i, consults D to make sure the edge connecting to the vertex at index i+1 is non-zero
    if(D[z[i],z[i+1]] == 0)
      return(0)
  }
  return(1)
}

distance_travelled <- function(L){
  C <- unlist(L)
  f <- D[C[20],C[1]] #endpoints require special care
  for(i in 1:19){
    f = f + D[C[i],C[i+1]]  #similar to indicatorPermissible, except it tallys the edges
  }
  return(f)
}

tspMCMC <- function(n,m){
v <- sample(U,size = 20, replace=FALSE) #without repetition sample first sequence
while(indicatorPermissible(v)==0){#make sure its valid
  v <- sample(U,size=20, replace=FALSE) 
}
S <- list(v) #record it in the state space
rejection_counter <- 0
trials <- 1
i <- 1
j <- 1
repetition_counter <- 0
travel_time <- rep(0, m+1) #init. the travel time vector
while(i <= m+n){
  P <- sample(U,size=1) #basic sampling 
  Q <- sample(U,size=1) #likewise for proposed state, 
  index_p = match(P, unlist(S[i])) #the position of P in the sequence
  index_q = match(Q, unlist(S[i])) #likewise for the second element in the pair
  k <- unlist(S[i])  
  z <- c(replace(k, c(index_p,index_q), c(Q,P))) #interchange P and Q 
  if(indicatorPermissible(z) == 1){#if the new sequence is valid
    if(i >= n){#has equilibrium been reached yet?
      travel_time[j] = distance_travelled(S[i]) #store travel time
      j <- j + 1
    }
    S[i+1] <- list(z) #adds the new sequence to collection of sampled state spaces
    i <- i + 1 
  }
  else{
    if(i<=n){#this is done to make sure we dont double count travel times
      S[i+1] <- S[i]
      i <- i + 1
    }
    rejection_counter = rejection_counter+1 #see how many time our proposed transition introduced forbidden travel
  }
}
   answer <- mean(travel_time)
   cat("tspMCM(",n,",",m,")", "\n")
   cat("Rejection Counter: ", rejection_counter, "\n")
   cat("Average Return Time:", answer, "units" , "\n")
}
start <- Sys.time()
tspMCMC(1000,5000)
tspMCMC(2000,10000)
tspMCMC(4000,20000)
tspMCMC(60000,300000)
cat("Computational Length:", Sys.time()-start, "seconds", "\n")
```

### Appendix 

Below is the code. 
<!-- <a href="https://raw.githubusercontent.com/Kofi114/TravellingSalesman/main/task4/circuit.R" download> Code </a> -->
<!---Interactive-->

<iframe style="margin:0 auto; min-width: 100%;" id="circuitFrame" class="interactive" src="https://blackmonk.shinyapps.io/CircuitR/" scrolling="no" frameborder="no"></iframe>

<!---------------->
<!-- The following R file contains our <a href="/task4/circuit.R" download> code </a> -->
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
<script>
  iFrameResize({}, ".interactive");
</script>
