---
title: "Metropolis Algorithm for Expectations"
output: 
  html_document:
    theme: readable
includes:
      after_body: footer.html
---

```{r include= FALSE}
library(igraph)

```
The objective of this website is do project is to use a Metropolis algorithm with $\mu=1$(no rejection) to simulate the expected travel time for a Hamiltonian circuit chosen uniformly. A Hamiltonian circuit is defined as a sequence which starts at a predefined vertex, visits every other vertex in the state space once and returns to its start point. The algorithm shines when its impossible to sample from $\mu$,


There will be five pair of cities for which travel is forbidden. Since there are about ~$20!$ possible circuits(see visual below) the generator is constructed by choosing 2 cities uniformly from c(1:20) from a valid sequence and interchanging them to create a brand new sequence given the proposed sequence granted it is also valid.  
```{r echo=FALSE, webgl=TRUE, fig.width=12, fig.height=12, fig.align="center", fig.cap="2-D Graph with 20 vertices and 185 edges."}
D <- matrix(runif(n=20*20, min=0, max =1), ncol = 20) #uniformly distributed distance matrix
D[4,6] = 0
D[9,13]= 0
D[1,19]= 0 
D[10,17]= 0  
D[3,8] = 0 #NO TRAVEL NO DISTANCE
D[lower.tri(D)] = t(D)[lower.tri(D)] #symmetric matrix
diag(D) <- rep(0,20) #diagonals are 0

g <- graph_from_adjacency_matrix(D, mode="undirected" , weighted=TRUE)
plot(g, vertex.color="red")

```

