---
title: "Travelling Salesman"
output: 
  html_document:
    theme: readable
includes:
      after_body: footer.html
---

```{r include= FALSE}
library(igraph)

```
The objective of our project is to use a Metropolis algorithm with $\mu=1$ to simulate the expected travel time for a Hamiltonian circuit chosen uniformly. A Hamiltonian circuit is defined as a sequence which starts at a predefined vertex, visits every other vertex in the state space once and returns to its start point.


There will be five pair of cities for which travel is forbidden. Since there are about ~$20!$ possible circuits(see visual below) the generator is constructed by choosing 2 cities uniformly from c(1:20) from a valid sequence and interchanging them to create a brand new sequence given the proposed sequence granted it is also valid.  
```{r echo=FALSE, webgl=TRUE, fig.width=12, fig.height=12}
D <- matrix(runif(n=20*20, min=0, max =1), ncol = 20) #uniformly distributed distance matrix
D[4,6] = 0
D[9,13]= 0
D[1,19]= 0 
D[10,17]= 0  
D[3,8] = 0 #NO TRAVEL NO DISTANCE
D[lower.tri(D)] = t(D)[lower.tri(D)] #symmetric matrix
diag(D) <- rep(0,20) #diagonals are 0

g <- graph_from_adjacency_matrix(D, mode="undirected" , weighted=TRUE)
plot(g, vertex.color="red")

```

